---
title: "Blackrock Waterfowl Management Area"
subtitle: "Flooded acreage monitoring dashboard"
format:
  html:
    code-tools: false
    toc: false
---

```{=html}
<div class="dashboard-hero">
  <div class="hero-row">
    <div class="hero-main">
      <div class="hero-tag">Sentinel-2 · BWMA · Owens Valley</div>
      <h1>Blackrock Waterfowl Management Area</h1>
      <p class="hero-lead">Monthly flooded acreage monitoring · Inyo County, California</p>
    </div>
    <div class="hero-meta">
      <span>NIR threshold</span>
      <span>15-day composite</span>
    </div>
  </div>
</div>
```

```{r read_csvs, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(stringr)
library(lubridate)

csv_dir <- "flood_reports/csv_output"

csv_files <- list.files(
  path       = csv_dir,
  pattern    = "^flood_report_data_[0-9]{4}-[0-9]{2}-[0-9]{2}_.*\\.csv$",
  full.names = TRUE
)

monthly_raw <- lapply(csv_files, function(path) {
  fname     <- basename(path)
  date_part <- str_match(
    fname,
    "^flood_report_data_([0-9]{4}-[0-9]{2}-[0-9]{2})_.*\\.csv$"
  )[, 2]
  report_dt <- as.Date(date_part, format = "%Y-%m-%d")

  read_csv(path, show_col_types = FALSE) %>%
    mutate(report_date = report_dt)
}) %>%
  bind_rows() %>%
  select(Flood_Unit, acres_flooded, report_date)
```

```{r dashboard_stats, echo=FALSE, message=FALSE, warning=FALSE}
#| classes: stat-chunk
library(dplyr)
library(lubridate)
library(htmltools)

fmt_acres <- function(x) format(round(x, 1), nsmall = 1, trim = TRUE)

acres_value <- function(x, unit = "ac") {
  tags$span(
    class = "stat-value",
    style = "font-size:1.15rem;font-weight:700;line-height:1.15;white-space:nowrap;",
    tags$span(class = "stat-acres-num", style = "font-family:Georgia,Cambria,serif;font-variant-numeric:tabular-nums;font-size:1.25rem;color:#0b3d91;", x),
    tags$span(class = "stat-acres-unit", style = "font-family:ui-monospace,monospace;font-size:0.78rem;font-weight:600;color:#5a6578;margin-left:0.1rem;", unit)
  )
}

stat_card_style <- "background:#f8f9fb;border:1px solid #c5cdd8;border-left:3px solid #0b3d91;border-radius:3px;padding:0.5rem 0.6rem;"
stat_highlight_style <- paste0(stat_card_style, "border-left-color:#fc3d21;background:#fff;")
stat_grid_style <- "display:grid;grid-template-columns:repeat(3,1fr);gap:0.45rem;margin-bottom:0.45rem;"
stat_units_style <- "display:flex;flex-wrap:wrap;align-items:baseline;gap:0.35rem 0.65rem;margin-bottom:1rem;padding:0.45rem 0.6rem;background:#f8f9fb;border:1px solid #c5cdd8;border-left:3px solid #0b3d91;border-radius:3px;"
stat_label_style <- "font-family:ui-monospace,monospace;font-size:0.58rem;font-weight:600;letter-spacing:0.1em;text-transform:uppercase;color:#5a6578;margin-bottom:0.2rem;"
stat_detail_style <- "margin-top:0.25rem;font-size:0.68rem;color:#5a6578;line-height:1.25;"
stat_units_list_style <- "font-family:ui-monospace,monospace;font-size:0.68rem;color:#5a6578;line-height:1.35;"

totals <- monthly_raw %>%
  filter(Flood_Unit == "Total") %>%
  arrange(desc(report_date))

latest <- totals %>% slice(1)
prev   <- totals %>% slice(2)

latest_month <- floor_date(latest$report_date, "month")
season_peak <- monthly_raw %>%
  filter(Flood_Unit == "Total", report_date >= as.Date("2024-09-01")) %>%
  mutate(month_date = floor_date(report_date, "month")) %>%
  filter(month(report_date) %in% c(11, 12, 1, 2, 3)) %>%
  group_by(month_date) %>%
  slice_max(report_date, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  slice_max(acres_flooded, n = 1, with_ties = FALSE)

delta_detail <- if (nrow(prev) == 1) {
  delta <- latest$acres_flooded - prev$acres_flooded
  cls <- if (delta >= 0) "stat-delta-up" else "stat-delta-down"
  sign <- if (delta >= 0) "+" else ""
  tags$span(
    class = cls,
    paste0(sign, fmt_acres(delta), " ac vs prior month")
  )
} else {
  tags$span("First report in series")
}

season_value <- if (nrow(season_peak) == 1) {
  tags$span(
    acres_value(fmt_acres(season_peak$acres_flooded)),
    tags$span(
      class = "stat-acres-unit",
      style = "font-family:ui-monospace,monospace;font-size:0.78rem;font-weight:600;color:#5a6578;margin-left:0.25rem;",
      paste0("in ", format(season_peak$month_date[1], "%b %Y"))
    )
  )
} else {
  tags$span(class = "stat-value", "—")
}

mom_value <- if (nrow(prev) == 1) {
  acres_value(fmt_acres(abs(latest$acres_flooded - prev$acres_flooded)), "ac change")
} else {
  tags$span(class = "stat-value", "—")
}

tagList(
  tags$div(
    class = "stat-grid",
    style = stat_grid_style,
    tags$div(
      class = "stat-card stat-highlight",
      style = stat_highlight_style,
      tags$div(class = "stat-label", style = stat_label_style, "Latest flooded"),
      acres_value(fmt_acres(latest$acres_flooded)),
      tags$div(
        class = "stat-detail",
        style = paste0(stat_detail_style, "white-space:nowrap;"),
        paste0(format(latest_month, "%b %Y"), " · 15-day composite")
      )
    ),
    tags$div(
      class = "stat-card",
      style = stat_card_style,
      tags$div(class = "stat-label", style = stat_label_style, "Month-over-month change"),
      mom_value,
      tags$div(class = "stat-detail", style = stat_detail_style, delta_detail)
    ),
    tags$div(
      class = "stat-card",
      style = stat_card_style,
      tags$div(class = "stat-label", style = stat_label_style, "Season peak"),
      if (nrow(season_peak) == 1) season_value else tags$span(class = "stat-value", "—"),
      tags$div(class = "stat-detail", style = stat_detail_style, "Nov–Mar · 500 ac target")
    )
  ),
  tags$div(
    class = "stat-units-bar",
    style = stat_units_style,
    tags$span(class = "stat-label", style = paste0(stat_label_style, "margin-bottom:0;"), "Reporting units (7)"),
    tags$span(
      class = "stat-units-list",
      style = stat_units_list_style,
      "Drew · Waggoner · West Winterton · East Winterton · South Winterton · Thibaut Ponds · Thibaut"
    )
  )
)
```

::: {.dashboard-section}
::: {.dashboard-section-header}
## Monthly flooded acreage by unit

Stacked totals by flood unit. Dashed line marks the **500 ac** seasonal target (Nov-Mar).
:::

::: {.dashboard-card .chart-card}
```{r monthly_flooded_acres, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.width=10, fig.height=5.5, fig.align='center'}
library(dplyr)
library(lubridate)
library(ggplot2)

unit_order <- c(
  "Drew", "Waggoner", "West Winterton", "East Winterton",
  "South Winterton", "Thibaut Ponds", "Thibaut"
)

unit_colors <- c(
  "Drew"            = "#2e86ab",
  "Waggoner"        = "#e85d04",
  "West Winterton"  = "#7eb8a4",
  "East Winterton"  = "#a8c686",
  "South Winterton" = "#c4a35a",
  "Thibaut Ponds"   = "#8b6fad",
  "Thibaut"         = "#1a5f7a"
)

monthly_fixed <- monthly_raw %>%
  mutate(
    Flood_Unit = recode(Flood_Unit,
      "Waggonner" = "Waggoner",
      .default    = Flood_Unit
    ),
    month_date = floor_date(report_date, unit = "month")
  )

monthly_avg <- monthly_fixed %>%
  filter(Flood_Unit != "Total") %>%
  filter(month_date >= as.Date("2024-09-01")) %>%
  group_by(Flood_Unit, month_date) %>%
  slice_max(report_date, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(Flood_Unit, month_date, acres_flooded) %>%
  rename(avg_acres = acres_flooded) %>%
  mutate(Flood_Unit = factor(Flood_Unit, levels = unit_order))

x_min <- min(monthly_avg$month_date) - 15
x_max <- max(monthly_avg$month_date) + 15

monthly_avg %>%
  ggplot(aes(x = month_date, y = avg_acres, fill = Flood_Unit)) +
  geom_col(width = 22, color = "white", linewidth = 0.15) +
  geom_hline(
    yintercept = 500,
    linetype   = "dashed",
    color      = "#b8860b",
    linewidth  = 0.65
  ) +
  annotate(
    "label",
    x     = x_max - 5,
    y     = 500,
    label = "500 ac target (Nov-Mar)",
    hjust = 1,
    vjust = -0.35,
    size  = 2.8,
    fontface = "bold",
    color = "#7a5c10",
    fill  = "#fffef8",
    label.size    = 0.3,
    label.padding = grid::unit(0.28, "lines")
  ) +
  scale_fill_manual(values = unit_colors, breaks = unit_order, name = "Flood unit") +
  scale_x_date(
    limits      = c(x_min, x_max),
    date_breaks = "1 month",
    date_labels = "%b '%y",
    expand      = c(0.02, 0)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.06)), labels = scales::comma) +
  coord_cartesian(clip = "off") +
  labs(
    title = NULL,
    x     = NULL,
    y     = "Acres flooded"
  ) +
  theme_minimal(base_size = 13, base_family = "sans") +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_line(color = "#e8eef2", linewidth = 0.4),
    axis.text.x        = element_text(angle = 45, hjust = 1, color = "#5c6b73"),
    axis.text.y        = element_text(color = "#5c6b73"),
    axis.title.y       = element_text(color = "#1f2933", face = "bold"),
    legend.position    = "bottom",
    legend.title       = element_text(face = "bold", size = 10),
    legend.text        = element_text(size = 9),
    legend.key.size    = unit(0.45, "cm"),
    plot.margin        = margin(8, 12, 8, 8)
  )
```
:::
:::

::: {.dashboard-section}
::: {.dashboard-section-header}
## Monthly report archive

Each cell links to that month’s archived report. The **day number** is the **file label date** (first scene in the collection)—not a single-scene acquisition. **Since April 2026**, reports follow the composite workflow below; older entries may reflect ad-hoc run dates and fewer scenes in the window.
:::

::: {.dashboard-card .table-card}
```{r monthly_raw, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(lubridate)
library(tidyr)
library(knitr)
library(stringr)

report_scan_dir <- "flood_reports/reports"
report_link_dir <- "flood_reports/reports"

html_files <- list.files(
  path       = report_scan_dir,
  pattern    = "^bwma_flood_report_[0-9]{4}-[0-9]{2}-[0-9]{2}_.*\\.html$",
  full.names = FALSE
)

link_style <- "font-family:ui-monospace,monospace;font-size:0.75rem;font-weight:600;white-space:nowrap;text-decoration:underline;color:#0b3d91;"

df_links <- tibble(file = html_files) %>%
  mutate(
    date       = as.Date(str_extract(file, "[0-9]{4}-[0-9]{2}-[0-9]{2}")),
    year       = year(date),
    month_num  = month(date),
    month_name = month(date, label = TRUE, abbr = TRUE),
    fy_num     = if_else(month_num >= 7, year, year - 1L),
    fy_label   = paste0(
      fy_num, "-",
      str_sub(as.character(fy_num + 1L), 3, 4)
    ),
    day_only   = format(date, "%d"),
    link_title = if_else(
      date >= as.Date("2026-04-01"),
      paste0("Composite report · label date ", format(date, "%b %d, %Y")),
      paste0("Archived report · label date ", format(date, "%b %d, %Y"), " (pre-composite workflow)")
    ),
    full_link  = paste0(
      "<a href='", report_link_dir, "/", file,
      "' target='_blank' title='", link_title,
      "' style='", link_style, "'>", day_only, "</a>"
    )
  )

df_one_per_month <- df_links %>%
  group_by(fy_label, month_name) %>%
  slice_min(date, n = 1, with_ties = FALSE) %>%
  ungroup()

all_months_ord <- ordered(month.abb, levels = month.abb)

wide_tbl <- df_one_per_month %>%
  select(fy_label, month_name, full_link) %>%
  complete(
    fy_label,
    month_name = all_months_ord,
    fill       = list(full_link = "")
  ) %>%
  mutate(
    month_for_fy = factor(
      as.character(month_name),
      levels = c(
        "Jul", "Aug", "Sep", "Oct", "Nov", "Dec",
        "Jan", "Feb", "Mar", "Apr", "May", "Jun"
      )
    )
  ) %>%
  select(fy_label, month_for_fy, full_link) %>%
  pivot_wider(
    names_from   = month_for_fy,
    values_from  = full_link,
    names_expand = TRUE
  ) %>%
  arrange(fy_label) %>%
  select(fy_label, any_of(c("Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun"))) %>%
  rename(`Year` = fy_label)

knitr::kable(
  wide_tbl,
  format     = "html",
  escape     = FALSE,
  caption    = "Monthly Reports",
  table.attr = "class='table dashboard-reports-table'"
)
```

<p class="table-legend">Click a day number to open the HTML report (unit table, map, downloads).</p>

```{=html}
<div class="table-archive-note">
  <strong>Composite workflow (Apr 2026 onward):</strong>
  Run mid-month with start date <code>YYYY-MM-01</code>. All qualifying Sentinel-2 scenes in the 15-day window are cloud-masked and combined into a <strong>median composite</strong>; acreage uses NIR on that composite (threshold 0.14).
  <br><br>
  <strong>Earlier archive (before Apr 2026):</strong>
  The script always built a median composite, but reports were often run on ad-hoc dates when a clear scene was available—sometimes with only one pass in the window. Legacy report pages may say “imagery date”; that label is the filename date, not proof of a single scene.
</div>
```
:::
:::

```{=html}
<!-- Latest monthly run: July 2026 (threshold 0.14). June: flood_reports/reports/bwma_flood_report_2026-06-01_0.14.html -->
<div class="dashboard-section">
  <div class="report-embed-card">
    <div class="report-embed-header">
      <div>
        <h2>Most recent report</h2>
        <p>July 2026 · Sentinel-2 composite · threshold 0.14</p>
      </div>
      <a href="flood_reports/reports/bwma_flood_report_2026-07-01_0.14.html" target="_blank" rel="noopener">Open full report ↗</a>
    </div>
    <iframe
      src="flood_reports/reports/bwma_flood_report_2026-07-01_0.14.html"
      class="report-iframe"
      id="report-frame"
      title="Latest BWMA flood acreage report">
    </iframe>
  </div>
</div>
```
